---
title: "Random Variables"
date: "2023-06-28"
categories: [statistics]
format:
  html: 
    code-fold: false
---

## Rnadom Variables

$\textbf{Definition 1:}$ A random variable is a function $X$ which assigns to each element $s \in S$ one and only one number $X(s) = x$, where $x \in S_{x} \subseteq \mathbb{R}$, i.e.,
$$
X: S \longrightarrow S_{x} \subseteq \mathbb{R}
$$
Random variables help us to view the collected data as an *outcome* of a random experiment, which is a procedure whose outcome is unknown to the experimenter before it takes place.

In practice, we often use $\Omega_{X}$ to denote support, or say sample sample, of random variable $X$. However, there are numerous notations to denote a support, e.g., $\Omega, S, S_{X}, \operatorname{supp}(X)$.

Take throwing coins for example, if we only throw the coin for one time, then the sample space would be $\{H, T\}$, where $H$ represents head and $T$ represents tail. If we throw the coin for two times, now the sample space would be $\{HH, HT, TH, TT\}$.

Another example is tossing two dice. We toss two dice. In this case, the sample space is
$$
\begin{aligned}
S = \{(1, 1), (1, 2), (1, 3), (1, 4), (1, 5), (1, 6),\\
(2, 1), (2, 2), (2, 3), (2, 4), (2, 5), (2, 6),\\
(3, 1), (3, 2), (3, 3), (3, 4), (3, 5), (3, 6),\\
(4, 1), (4, 2), (4, 3), (4, 4), (4, 5), (4, 6),\\
(5, 1), (5, 2), (5, 3), (5, 4), (5, 5), (5, 6),\\
(6, 1), (6, 2), (6, 3), (6, 4), (6, 5), (6, 6)\}
\end{aligned}
$$

One may define the random variable X as the sum:
$$
S_{X} = \{2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12\}.
$$
or let the random variable $Y$ be the product:
$$
S_{Y} = \{1, 2, 3, 4, 5, 6, 8, 9, 10, 12, 15, 16, 18, 20, 24, 25, 30, 36\}.
$$

We can use Monte Carlo experiment to simlulate the process and result of tossing a dice.

```{r}
# d for dice
d <- sample(1:6, size = 100, replace = TRUE)
d
```

Using `table` function we can see the probabilty of each outcome when tossing a dice.

```{r}
table(d) / 100
plot(table(d) / 100)

# Add a title
title("Probability Mass Function (pmf)")
```

If we toss it for 1000000 times, then we will find that the probability of each outcome would be very close to $1/6 \approx 0.166667$, which yields a conclusion that given a larger the sample size, the closer the relative frequencies to the probabilities would be.

```{r}
d <- sample(1:6, size = 1000000, replace = TRUE)
table(d) / 1000000
```

Note that we can use the support of to categorize random variable. For a discrete random variable, its support must be a discrete set, while for a continuous random variable, its support is an interval of values.


### Discrete Random Variables

$\textbf{Definition 2:}$ If a random variable $X$ has a discrete distribution with $\operatorname{supp}(X) \in \mathbb{R}$, and define a function $f(x): \mathbb{R} \mapsto [0,1]$ to be
$$
f(x) = 
\begin{cases}
  \mathbb{P}(X = x),& \quad x \in \operatorname{supp}(X)\\
  0,& \quad x \notin \operatorname{supp}(X)
\end{cases}
$$
which satisfies:

- $f(x) > 0, \forall x \in \operatorname{supp}(X)$
- $\sum_{x \in \operatorname{supp}(X)} f(x) = 1$ 
- $\mathbb{P}(X \in A) = \sum_{x \in A}f(x)$, where $A \in \operatorname{supp}(X)$

then we say $f(x)$ is a *probability mass function* (pmf).

$\textbf{Definition 3:}$ We say a function $F(x)$ is a *cumulative distribution function* (CDF) if for any $x \in \mathbb{R}$, $F(x): \mathbb{R} \mapsto [0, 1]$, which satisfies
$$
F(x) = \mathbb{P}(X \leq x) = \sum_{x_{i} \in x}f(x)
$$
Here are some properties of discrete cumulative distribution function:

1. $\mathbb{P}(X > a) = 1 - \mathbb{P}(X \leq a) = 1 - F(a)$
2. $\mathbb{P}(X \geq a) = \mathbb{P}(X = a) + \mathbb{P}(X > a) = f(a) + 1 - F(a) = [1 - F(a)] + f(a)$
3. $\mathbb{P}(a < X \leq b) = F(b) - F(a)$
4. $\mathbb{P}(a < X < b) = F(b) - F(a) - f(b)$
5. $\mathbb{P}(a \leq X \leq b) = F(b) - F(a) + f(a)$
6. $\mathbb{P}(a \leq X < b) = F(b) - F(a) + f(a) - f(b)$
7. $\lim_{x \to -\infty}F(x) = 0$, and $\lim_{x \to \infty}F(x) = 1$

### Continuous Random Variables

$\textbf{Definition 4:}$ For a continuous random variable $X$, if $f(x): \mathbb{R} \mapsto \mathbb{R}$ and for any real number $a \leq b$ satisfy
$$
\mathbb{P}(a < X \leq b) = \int^{b}_{a}f(x) dx = F(x)\Big|^{b}_{a} = F(b) - F(a)
$$
and

- $f(x) > 0, \forall \operatorname{supp}(X) \in \Omega_{X}$
- $\int_{x \in \Omega_{X}}f(x)dx = 1$

then we say $f(x)$ is a *probability density function* (PDF) of $X$.

```{r}
# Generate x values for the range of interest
x <- seq(-4, 4, 0.01)

# Compute corresponding y values using the standard normal density function
y <- dnorm(x, mean = 0, sd = 1)

# Create a blank plot with the desired x-axis limits and y-axis labels
plot(x, y, type = "l", xlab = "x", ylab = "Density", xlim = c(-4, 4), ylim = c(0, 0.45), axes = F)

# Fill the region between -1 and 1 with gray color
polygon(c(-1, x[x >= -1 & x <= 1], 1), c(0, y[x >= -1 & x <= 1], 0), col = "gray")

axis(1)
axis(2)

# Add a title
title("Probability Density Function (PDF)")
```
Note that there is a corr

```{r, fig.align='center'}
# Generate x values for the range of interest
x <- seq(-4, 4, 0.01)

# Compute corresponding y values for PDF and CDF
PDF <- dnorm(x, mean = 0, sd = 1)
CDF <- pnorm(x, mean = 0, sd = 1)

# Set up a layout with two plots side by side
layout(matrix(c(1, 2), nrow = 1, ncol = 2), widths = c(0.5, 0.5))

# Plot PDF on the left
par(mar = c(5, 4, 4, 2))
plot(x, PDF, type = "l", xlab = "x", ylab = "Density", main = "PDF")
abline(v = c(-1, 1), lty = 2, col = "gray")  # Add vertical lines at -1 and 1

# Plot CDF on the right
par(mar = c(5, 2, 4, 4))
plot(x, CDF, type = "l", xlab = "x", ylab = "Probability", main = "CDF")
abline(v = c(-1, 1), lty = 2, col = "gray")  # Add vertical lines at -1 and 1
```
Mathematically, by using the fundamental theorem of calculus, we can find the relationship. Given a well-defined random variable $X$. Write its CDF as $F(x)$ and its PDF as $f(x)$, Since
$$
\int^{b}_{a} f(x) dx = F(x) \Big|^{b}_{a}
$$
then
$$
f(x) = \frac{\partial F(x)}{\partial x}, \quad F(x) = \int^{x}_{-\infty}f(u)du.
$$
Now we will discuss some properties of CDF given continuous random variable. Suppose $X$ is a continuous random variable, and $\operatorname{supp}(X) = \{x| -\infty < x < \infty\} = \{x | x \in \mathbb{R}\}$. Then

1. $F(x) = \mathbb{P}(X \leq x) = \int^{x}_{-\infty}f(u)du = \mathbb{P}(X < x)$
2. If $a < b$, then $F(a) \leq F(b)$
3. $\mathbb{P}(a < X < b) = \mathbb{P}(a \leq X \leq b) = F(b) - F(a)$
4. $\lim_{x \to -\infty}F(x) = 0$, and $\lim_{x \to \infty}F(x) = 1$
5. $\mathbb{P}(X \geq x) = \mathbb{P}(X > x) = 1- \mathbb{P}(X \leq x) = 1- F(x)$

